---
title: "hw-2"
output: github_document
date: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## data

The first step in completing this problem set requires me to reformat the provided data file, `hw02_ltdb_2010.dta`, into a csv that capable of being read and mutated in RStudio.

To accomplish this, we'll need to load a few packages.

```{r}
library(haven)
library(tidyverse)
library(dplyr)
```

Now, we can load the data into our environment: 

```{r}
hw02_ltdb_2010 <- read_dta("~/Downloads/hw02_ltdb_2010.dta")
```

## q1

The first question requires us to select one of the 100 largest CBSAs given in the dataset, which is provided by the cbsasamp1 variable. 

I chose the Milwaukee-Waukesha-West Allis, WI CBSA. Based on this decision, let's go ahead and mutate the dataframe that all rows containing data outside this CBSA are removed.

```{r}
milwaukee_cbsa <- hw02_ltdb_2010 %>% filter(cbsaname == "Milwaukee-Waukesha-West Allis, WI CBSA")
```

Now, we have to generate three tables describing this CBSA: one to summarize the population characteristics, one to summarize relevant neighborhood attributes, and one to quantify the levels of neighborhood segregation.

### table 1

```{r}
pop_characteristics <- read_csv("/Users/kenjinchang/github/pam6950/pop_characteristics.csv")
```

```{r}
print(as_tibble(pop_characteristics))
```
### table 2

```{r}
neighbor_attrib <- read_csv("/Users/kenjinchang/github/pam6950/neighbor_attrib.csv")
```

```{r}
print(as_tibble(neighbor_attrib))
```

### table 3

```{r}
neighbor_seg <- read_csv("/Users/kenjinchang/github/pam6950/neighbor_seg.csv")
```

```{r}
print(as_tibble(neighbor_seg))
```

## q2

For this portion, I've created a dedicated dataframe to capture the distribution in tract-level exposure index values, both for White- and Hispanic-identifying individuals in the Milwaukee CBSA and for poverty-stricken and non-poverty-stricken individuals.

```{r}
q2 <- read_csv("/Users/kenjinchang/github/pam6950/q2.csv")
```

```{r}
ggplot(q2,aes(x=whitexhispanic)) +
  geom_histogram(bins=75,color="gray50",size=.1,alpha=.8) +
  xlab("White-Hispanic Interaction") +
  ylab("Count") +
  theme_classic()
```
```{r}
ggplot(q2,aes(x=povxnopov)) +
  geom_histogram(bins=75,color="gray50",size=.1,alpha=.8) +
  xlab("Poverty-No Poverty Interaction") +
  ylab("Count") +
  theme_classic()
```

## q3

Still merging...

## q4

```{r}
library(choroplethr)
library(choroplethrZip)
library(ggplot2)
library(tmap)
library(tidycensus)
```

Experimenting with tidycensus:

```{r}
medincome <- get_acs(
  geography="tract",
  variables="B19013_001",
  state="WI",
  geometry=TRUE,
  year=2020)
plot(medincome["estimate"])
```

Experimenting with ggplot2:

```{r}
ggplot(medincome,aes(fill=estimate)) + 
  geom_sf() + 
  scale_fill_distiller(palette="RdPu",
                       direction=1) + 
  labs(title="Median Household Income by Tract, 2020",
       caption="Data Source: 2015-2020 ACS, US Census Bureau",
       fill="ACS Estimate") + 
  theme_void()
```

