---
title: "hw-2"
output: github_document
date: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## data

The first step in completing this problem set requires me to reformat the provided data file, `hw02_ltdb_2010.dta`, into a csv that capable of being read and mutated in RStudio.

To accomplish this, we'll need to load a few packages.

```{r}
library(haven)
library(tidyverse)
library(dplyr)
```

Now, we can load the data into our environment: 

```{r}
hw02_ltdb_2010 <- read_dta("~/Downloads/hw02_ltdb_2010.dta")
```

## q1

The first question requires us to select one of the 100 largest CBSAs given in the dataset, which is provided by the cbsasamp1 variable. 

I chose the Milwaukee-Waukesha-West Allis, WI CBSA. Based on this decision, let's go ahead and mutate the dataframe that all rows containing data outside this CBSA are removed.

```{r}
milwaukee_cbsa <- hw02_ltdb_2010 %>% filter(cbsaname == "Milwaukee-Waukesha-West Allis, WI CBSA")
```

Now, we have to generate three tables describing this CBSA: one to summarize the population characteristics, one to summarize relevant neighborhood attributes, and one to quantify the levels of neighborhood segregation.

### table 1

```{r}
pop_characteristics <- read_csv("/Users/kenjinchang/github/pam6950/pop_characteristics.csv")
```

```{r}
print(as_tibble(pop_characteristics))
```
### table 2

```{r}
neighbor_attrib <- read_csv("/Users/kenjinchang/github/pam6950/neighbor_attrib.csv")
```

```{r}
print(as_tibble(neighbor_attrib))
```

### table 3

```{r}
neighbor_seg <- read_csv("/Users/kenjinchang/github/pam6950/neighbor_seg.csv")
```

```{r}
print(as_tibble(neighbor_seg))
```

## q2

