---
title: "hw-5"
output: github_document
date: "2023-03-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## data prep and package loading

after importing the `tomp_poverty.dta` file into our environment, we can store the block group poverty data for tompkins country as a csv in our class repo using the following: 

```{r,eval=FALSE}
write.csv(tomp_poverty,"tomp_poverty.csv")
print(tomp_poverty)
```

```{r}
library(tidyverse)
```


## q1

### a 

first, we need to generate a new variable aggregating the number of families above and below poverty with children aged 0-17.

```{r}
tomp_poverty <- read.csv("tomp_poverty.csv")
```


```{r}
tomp_poverty %>% 
  mutate(fkids=fpovkids+fnopovkids)
```

with this new variable in place, we can then construct another dividing the number of families with children below poverty by the total number of families with children.

```{r}
tomp_poverty %>% 
  mutate(fkids=fpovkids+fnopovkids,
         prop_fkids_pov=fpovkids/fkids)
```

we can then group by urbanicity to see how the proportions for each block group vary by designation:

```{r}
tomp_poverty %>% 
  mutate(fkids=fpovkids+fnopovkids,
         prop_fkids_pov=fpovkids/fkids) %>%
  group_by(urban) %>%
  summarize(povrate_fkids=mean(prop_fkids_pov,na.rm=TRUE))
```

based on these results, we can see that the mean poverty rate among families with children is greater in rural areas (0.155 [3sf]) than urban areas (0.105 [3sf]). however, from my one-tailed t-test, this difference appears to be insignificant (p=0.115[3sf]). poverty in tompkins county therefore appears to impact a relatively equal share of rural and urban families with children.

we can run a similar analysis for all families (i.e., not just those with children aged 0-17):

```{r}
tomp_poverty %>% 
  mutate(prop_f_pov=(fpov+fpovkids)/fams) %>%
  group_by(urban) %>%
  summarize(povrate_f=mean(prop_f_pov,na.rm=TRUE)) 
```

again, we see that families living in rural areas tend to experience poverty at a greater rate than families living in urban areas, though the difference does not appear to be significant (p=0.749[3sf]).

### b

```{r}
tomp_poverty %>% 
  mutate(tot_fpovkids=sum(fpovkids),
         tot_fnopovkids=sum(fnopovkids),
         di=abs(fpovkids/tot_fpovkids - fnopovkids/tot_fnopovkids)) %>%
  summarize(dissim=sum(di)/2)
```

when comparing families with children aged 0-17 who are living below poverty and families with children aged 0-17 who are not, we observed a dissimilarity index of 0.495 (3sf). this suggests that approximately half of all families with children living below poverty would need to change residence to create a more uniform distribution across block groups. because the index rangrs from 0.0 to 1.0, signifying complete integration and complete segregation, respectively, this index value appears to suggest that theire is a medium level of segregation. 

### c 

acs data are subject to errors in sampling, and the resulting biases from these errors are magnified when we use these data to observed phenomena at small spatial scales.

in addition to this scaling issue, there are also issues in the reliability of acs survey results, with patterned variation occurring across different surveys.

## q2

loading in necessary packages:

```{r}
library(tidycensus)
options(tigris_use_cache=TRUE)
```

loading in point data:

```{r, eval=FALSE}
write.csv(tomp_sites,"tomp_sites.csv")
print(tomp_sites)
```


```{r}
tomp_sites <- read.csv("tomp_sites.csv")
```

### a

```{r}
tompkins <- get_acs(
  state = "NY",
  county = "Tompkins",
  geography = "cbg",
  variables = "B11003_003",
  geometry = TRUE,
  year = 2016
)
```


```{r}
ggplot() + 
  geom_sf(data=tompkins,aes(fill=estimate)) + 
  geom_point(data=tomp_sites,aes(x=xlon,y=ylat),color="white") +
  scale_fill_viridis_c(option="magma") +
  xlab("") + 
  ylab("") +
  labs(fill="Families",caption="White points indicate possible locations of child-care centers\n Projected using NAD 1983") +
  ggtitle("Number of families with children aged 0-17 in each block group \nliving below poverty in Tompkins county") + 
  theme(axis.ticks=element_blank(),axis.text=element_blank(),panel.background=element_rect(fill=NA))
```

### b 

```{r}
ggplot() + 
  geom_sf(data=tompkins,aes(fill=estimate)) + 
  geom_point(data=tomp_sites,aes(x=xlon,y=ylat),color="white") +
  scale_fill_distiller(palette="Greys") +
  xlab("") + 
  ylab("") +
  labs(fill="Families",caption="White points indicate possible locations of child-care centers\n Projected using NAD 1983") +
  ggtitle("Number of families with children aged 0-17 in each block group \nliving below poverty in Tompkins county") + 
  theme(axis.ticks=element_blank(),axis.text=element_blank(),panel.background=element_rect(fill=NA))
```

i'm not sure if anything is necessarily lost, but it becomes more difficult, at least for me, to distinguish between the point and area data. this could possibly be resolved by adding a black border to the white-filled points to add contrast in the areas where it is most needed (i.e., for the centers positioned in block groups with area data that is near the maximum or minimum).



### c
